---
title: "Limbus_epi_comp"
author: "JGASmits"
date: "8/10/2021"
output: html_document
---

```{r setup, include=FALSE}
library(Seurat)
require("devtools")
library('ggplot2')
library('plyr')
library('dplyr')
library('tidyr')
library('mvoutlier')
library('limma')
library('knitr')
library('SingleCellExperiment')
library("scater")
library("Seurat")
library("scran")
library("RColorBrewer")
library("plot3D")
library("stringr")
library("SAVER")
library("parallel")
library('progeny')
library("ComplexHeatmap") 
library("org.Hs.eg.db")
library("clusterProfiler") 
library("ggpubr")
library("circlize")
library("cowplot")
library('Seurat')
library('clustree')
library('tidyverse')
library('monocle')
library('SeuratWrappers')
```

```{r}
work_dir <- '/ceph/rimlsfnwi/data/moldevbio/zhou/jsmits/Publications/gene-regulatory-networks-in-epidermal-and-corneal-epithelia'
dateoftoday <- gsub("-", "", as.character(Sys.Date()))
resultsdir <- paste0(workdir, dateoftoday)
system(paste("mkdir -p ", resultsdir))

```


## R Markdown

```{r cars}
Object_Epi_filepath <- "/ceph/rimlsfnwi/data/moldevbio/zhou/jsmits/Epi_fate_specifiers/scRNAseq/Epidermis_Ward.rds"
Object_limbus_filepath <- "/ceph/rimlsfnwi/data/moldevbio/zhou/jsmits/Epi_fate_specifiers/scRNAseq/lakoRNAannotated.rds"
```

```{r}
Epi_seurat <- readRDS(Object_Epi_filepath)
Epi_seurat = subset(Epi_seurat, subset = disease == "HC")
Idents(Epi_seurat) = Epi_seurat$annotation
Epi_seurat$dataset <- Epi_seurat$sample
Limbus_seurat <- readRDS(Object_limbus_filepath)
Idents(Limbus_seurat) = Limbus_seurat$costum_clustering
Limbus_seurat$annotation = Limbus_seurat$costum_clustering
Limbus_seurat$dataset <- Limbus_seurat$orig.ident
```

```{r}
pdf(paste0(resultsdir,'/atwood_Umap.pdf') ,width=6,height=6,paper='special')
DimPlot(Epi_seurat, label=FALSE)+theme(legend.position=c(0.85,0))
FeaturePlot(Epi_seurat, features = c("KRT14", "KRT5", "POSTN", "PTTG1",'FLG'))
dev.off()

LSC_subsets <- Limbus_seurat[,(Limbus_seurat$costum_clustering == 'LPCs')]

pdf(paste0(resultsdir,'/LAKO_Umap_PAX6_ELF.pdf') ,width=24,height=6,paper='special')
FeaturePlot(LSC_subsets, features = c("PAX6", "ELF3"), blend = TRUE)
FeatureScatter(LSC_subsets, feature1 = "PAX6", feature2 = "ELF3")
DoHeatmap(LSC_subsets, features = c('PAX6','ELF3')) 
dev.off()


pdf(paste0(resultsdir,'/LAKO_Umap.pdf') ,width=6,height=6,paper='special')
DimPlot(Limbus_seurat, label=FALSE)+theme(legend.position=c(0.85,0))
FeaturePlot(Limbus_seurat, features = c("S100A2", "PAX6", "TP63", "CPVL"))
FeaturePlot(Limbus_seurat, features = c("PAX6", "ELF3"), blend = TRUE)
dev.off()
```
```{r}
mat <- as.matrix(LSC_subsets@assays$RNA@scale.data[c('PAX6','ELF3'),])
#mat <- rbind(mat,LSC_subsets@meta.data$orig.ident)
mat <- mat[,order(mat[nrow(mat),])]

 grid.grabExpr(draw(Heatmap(mat, cluster_columns = T, cluster_rows = T,show_row_names = F,show_column_names = F)))

```

Lets generate pseudobulk for each population and each object
```{r}
# dataset <- 'LAKO'
# for (seur_obj in c(Limbus_seurat, Epi_seurat)){
#   print(seur_obj)
#   pseudobulk_df <- as.data.frame(row.names(seur_obj@assays$RNA@counts))
#   colnames(pseudobulk_df) <- 'gene'
#   for (cell_type in unique(Idents(seur_obj))){
#     print(cell_type)
#     pseudobulk_df[[cell_type]] <- rowSums(as.data.frame(seur_obj@assays$RNA@counts)[Idents(seur_obj)== cell_type])
#   }
#   row.names(pseudobulk_df) <- pseudobulk_df$gene
#   pseudobulk_df$gene <- NULL
#   
#   seur_obj <- NormalizeData(seur_obj, normalization.method = "LogNormalize", scale.factor = 10000)
#   seur_obj <- FindVariableFeatures(seur_obj, selection.method = 'vst', nfeatures = 5000)
#   variable_genes <- VariableFeatures(seur_obj)
#   
#   if (dataset == 'LAKO'){
#     write.table(pseudobulk_df, file = paste0(resultsdir,'/pseudobulk_Lako.tsv'), sep = '\t')
#     write.table(variable_genes, file = paste0(resultsdir,'/vargenes_Lako.tsv'), sep = '\t',row.names = F)
#     dataset <- 'Atwood'}
#   if (dataset == 'Atwood'){
#     write.table(pseudobulk_df, file = paste0(resultsdir,'/pseudobulk_Atwood.tsv'), sep = '\t')
#     write.table(variable_genes, file = paste0(resultsdir,'/vargenes_Atwood.tsv'), sep = '\t', row.names = F, col.names = F)
#                 }
# }
```

For the KCs lets generate pseudobulk inc seperate technical replica's, we will merge BASALI & BASALII KCs
Since the KCs are in multiple cell cycle phases, we will remove all cells that are not in G1 before generating pseudobulk.
For the LSCs lets generate pseudobulk inc seperate technical replica's, we will merge all epithelial clusters, LPCs, LNPCs, CB, CjS, and CSBs

```{r}
#lets first subset the seurat object to only contain the cell types of interests:
LPNCs_seurat <- Limbus_seurat[,(Limbus_seurat$costum_clustering == 'LPCs')]

#lets loop over the technical replica's and print how many cells each replica has
#let merge lako5,6,7&8 since those replicas combined contain only 400 LSCs cells#
# LPNCs_seurat$dataset <- gsub('lako8', 'lako2',LPNCs_seurat$dataset)
# LPNCs_seurat$dataset <- gsub('lako7', 'lako2',LPNCs_seurat$dataset)
# LPNCs_seurat$dataset <- gsub('lako6', 'lako2',LPNCs_seurat$dataset)
# LPNCs_seurat$dataset <- gsub('lako5', 'lako2',LPNCs_seurat$dataset)

#lets output all the replicas as pseudobulK:
pseudobulk_df <- as.data.frame(row.names(LPNCs_seurat@assays$RNA@counts))
colnames(pseudobulk_df) <- 'gene'

for (replica in unique(LPNCs_seurat$dataset)){
  print(replica)
  pseudobulk_df[[replica]] <- rowSums(as.data.frame(LPNCs_seurat@assays$RNA@counts)[LPNCs_seurat$dataset== replica])
}
row.names(pseudobulk_df) <- pseudobulk_df$gene
pseudobulk_df$gene <- NULL
write.table(pseudobulk_df, file = paste0(resultsdir,'/LPCs_replicas_pseudobulk_Lako.tsv'), sep = '\t')
```
all_ lako epithelial cells combined
```{r}
#lets first subset the seurat object to only contain the cell types of interests:
LPNCs_seurat <- Limbus_seurat[,(Limbus_seurat$costum_clustering == 'LPCs' | Limbus_seurat$costum_clustering == 'LNPCs' | Limbus_seurat$costum_clustering == 'CB'|Limbus_seurat$costum_clustering == 'CjS'|Limbus_seurat$costum_clustering == 'CSB')]

#lets loop over the technical replica's and print how many cells each replica has
#let merge lako5,6,7&8 since those replicas combined contain only 400 LSCs cells
# LPNCs_seurat$dataset <- gsub('lako8', 'lako2',LPNCs_seurat$dataset)
# LPNCs_seurat$dataset <- gsub('lako7', 'lako2',LPNCs_seurat$dataset)
# LPNCs_seurat$dataset <- gsub('lako6', 'lako2',LPNCs_seurat$dataset)
# LPNCs_seurat$dataset <- gsub('lako5', 'lako2',LPNCs_seurat$dataset)

#lets output all the replicas as pseudobulK:
pseudobulk_df <- as.data.frame(row.names(LPNCs_seurat@assays$RNA@counts))
colnames(pseudobulk_df) <- 'gene'

for (replica in unique(LPNCs_seurat$dataset)){
  print(replica)
  pseudobulk_df[[replica]] <- rowSums(as.data.frame(LPNCs_seurat@assays$RNA@counts)[LPNCs_seurat$dataset== replica])
}
row.names(pseudobulk_df) <- pseudobulk_df$gene
pseudobulk_df$gene <- NULL
write.table(pseudobulk_df, file = paste0(resultsdir,'/epithelial_replicas_pseudobulk_Lako.tsv'), sep = '\t')
```


```{r}
#lets first subset the seurat object to only contain the cell types of interests:
basalKCs_seurat <- Epi_seurat[,(Epi_seurat$annotation == 'BASAL I' | Epi_seurat$annotation == 'BASAL II' | Epi_seurat$annotation == 'BASAL III' )]

#Its all KCs with multiple stratification levels, lets take everyghint:
#basalKCs_seurat <- Epi_seurat
basalKCs_seurat <- CellCycleScoring(basalKCs_seurat, s.features = cc.genes.updated.2019$s.genes, g2m.features = cc.genes.updated.2019$g2m.genes, set.ident = TRUE)

#lets perform cell cycle scoring and remove all non G1 cells:
DimPlot(basalKCs_seurat, label=FALSE)+theme(legend.position=c(0.85,0))
DimPlot(Epi_seurat, label=FALSE)+theme(legend.position=c(0.85,0))

G1_KCs_seurat <- basalKCs_seurat[,basalKCs_seurat$Phase == 'G1']

#lets output all the replicas as pseudobulK:
pseudobulk_df <- as.data.frame(row.names(G1_KCs_seurat@assays$RNA@counts))
colnames(pseudobulk_df) <- 'gene'

for (replica in unique(G1_KCs_seurat$dataset)){
  print(replica)
  pseudobulk_df[[replica]] <- rowSums(as.data.frame(G1_KCs_seurat@assays$RNA@counts)[G1_KCs_seurat$dataset== replica])
}

row.names(pseudobulk_df) <- pseudobulk_df$gene
pseudobulk_df$gene <- NULL
write.table(pseudobulk_df, file = paste0(resultsdir,'/basal_replicas_pseudobulk_Atwood.tsv'), sep = '\t')
```
all_atwood epithelial cells combined

```{r}
#lets first subset the seurat object to only contain the cell types of interests:
basalKCs_seurat <- Epi_seurat
basalKCs_seurat <- CellCycleScoring(basalKCs_seurat, s.features = cc.genes.updated.2019$s.genes, g2m.features = cc.genes.updated.2019$g2m.genes, set.ident = TRUE)

#Its all KCs with multiple stratification levels, lets take everyghint:
basalKCs_seurat <- Epi_seurat
basalKCs_seurat <- CellCycleScoring(basalKCs_seurat, s.features = cc.genes.updated.2019$s.genes, g2m.features = cc.genes.updated.2019$g2m.genes, set.ident = TRUE)

#lets perform cell cycle scoring and remove all non G1 cells:
DimPlot(basalKCs_seurat, label=FALSE)+theme(legend.position=c(0.85,0))
DimPlot(Epi_seurat, label=FALSE)+theme(legend.position=c(0.85,0))

G1_KCs_seurat <- basalKCs_seurat[,basalKCs_seurat$Phase == 'G1']

#lets output all the replicas as pseudobulK:
pseudobulk_df <- as.data.frame(row.names(G1_KCs_seurat@assays$RNA@counts))
colnames(pseudobulk_df) <- 'gene'

for (replica in unique(G1_KCs_seurat$dataset)){
  print(replica)
  pseudobulk_df[[replica]] <- rowSums(as.data.frame(G1_KCs_seurat@assays$RNA@counts)[G1_KCs_seurat$dataset== replica])
}

row.names(pseudobulk_df) <- pseudobulk_df$gene
pseudobulk_df$gene <- NULL
write.table(pseudobulk_df, file = paste0(resultsdir,'/epithelial_replicas_pseudobulk_Atwood.tsv'), sep = '\t')
```



```{r}

pdf(paste0(resultsdir,'/atwood_Umap_cell_cycle.pdf') ,width=6,height=6,paper='special')
DimPlot(basalKCs_seurat, label=FALSE)+theme(legend.position=c(0.85,0))
DimPlot(Epi_seurat, label=FALSE)+theme(legend.position=c(0.85,0))
DimPlot(G1_KCs_seurat, label=FALSE)+theme(legend.position=c(0.85,0))
dev.off()

pdf(paste0(resultsdir,'/Lako_Umap_selection.pdf') ,width=6,height=6,paper='special')
DimPlot(LPNCs_seurat, label=FALSE)+theme(legend.position=c(0.85,0))
dev.off()
```


